---
title: "STATS 315B - Regressions"
author: "Nicolas Fertout"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Create Features

```{r}
require(knitr)
require(ggplot2)
require(dplyr)
require(tidyr)
```


```{r}
df_CGM_Char_HbA1c = read.csv("C:/Users/nicof/Stanford/STATS 315B/Project/df_CGM_Char_HbA1c.csv", row.names = NULL)
df_CGM_Char_HbA1c = df_CGM_Char_HbA1c[, -1]
df_feng = df_CGM_Char_HbA1c
```

```{r}
nrow(df_feng)
```


```{r}
# get only CGM columns for further analysis
df_feng_glucOnly = select(df_feng, starts_with("Glucose"))
```


```{r}
#compute rate of change (roc) of glucose. Since all obs are of the same \delta t, this is like taking the diff (scaled differently).
df_feng_roc = as.data.frame(t(apply(df_feng_glucOnly, 1, diff, na.rm = T)))
names(df_feng_roc) <- paste0(names(df_feng_roc), "_roc")
# compute statistics of roc
df_feng_roc_ready = df_feng_roc
df_feng_roc_ready$mean_gluc_roc <- rowMeans(df_feng_roc[, 1:(ncol(df_feng_roc))], na.rm = T)
df_feng_roc_ready$sd_gluc_roc <- apply(df_feng_roc, 1, function(x) {
  non_na_values <- x[!is.na(x)]
  if (length(non_na_values) > 1) {
    sd_val <- sd(non_na_values)
    ifelse(is.nan(sd_val) || is.na(sd_val), 0, sd_val)
  } else {
    0
  }
})
df_feng_roc_ready$min_gluc_roc <- apply(df_feng_roc, 1, min, na.rm = T)
df_feng_roc_ready$max_gluc_roc <- apply(df_feng_roc, 1, max, na.rm = T)
df_feng_roc_ready$median_gluc_roc <- apply(df_feng_roc, 1, median, na.rm = T)
# define helper function to get Median Absolute Deviation (MAD) of a df 
mad_fun <- function(x) {
  mad(x, na.rm = TRUE)
}
df_feng_roc_ready$mad_gluc_roc <- apply(df_feng_roc, 1, mad_fun)
```


```{r}
# df_feng = select(df_feng, -starts_with("PtID"))
# create new, not time dependent features
df_feng_stats = df_feng_glucOnly
df_feng_stats$mean_gluc <- rowMeans(df_feng_stats[, 1:(ncol(df_feng_stats))], na.rm = T)
df_feng_stats$sd_gluc <- apply(df_feng_glucOnly, 1, function(x) {
  non_na_values <- x[!is.na(x)]
  if (length(non_na_values) > 1) {
    sd_val <- sd(non_na_values)
    ifelse(is.nan(sd_val) || is.na(sd_val), 0, sd_val)
  } else {
    0
  }
})
df_feng_stats$min_gluc <- apply(df_feng_glucOnly, 1, min, na.rm = T)
df_feng_stats$max_gluc <- apply(df_feng_glucOnly, 1, max, na.rm = T)
df_feng_stats$median_gluc <- apply(df_feng_glucOnly, 1, median, na.rm = T)

# define helper function to get Median Absolute Deviation (MAD) of a df 
mad_fun <- function(x) {
  mad(x, na.rm = TRUE)
}
df_feng_stats$mad_gluc <- apply(df_feng_glucOnly, 1, mad_fun)

# Calculate the coefficient of variation
df_feng_stats$cv_gluc <- df_feng_stats$sd_gluc / df_feng_stats$mean_gluc * 100

# Calculate some proportions
df_feng_stats$prop_hypo <- apply(df_feng_glucOnly, 1, function(x) {
  non_na_values <- x[!is.na(x)]
  prop <- sum(x < 54, na.rm = TRUE) / length(non_na_values) 
  ifelse(is.nan(prop) || is.na(prop), 0, prop)
})

df_feng_stats$prop_clini_hypo <- apply(df_feng_glucOnly, 1, function(x) {
  non_na_values <- x[!is.na(x)]
  prop <- sum(x >= 54 & x <= 69, na.rm = TRUE) / length(non_na_values) 
  ifelse(is.nan(prop) || is.na(prop), 0, prop)
})

df_feng_stats$prop_target <- apply(df_feng_glucOnly, 1, function(x) {
  non_na_values <- x[!is.na(x)]
  prop <- sum(x >= 70 & x <= 180, na.rm = TRUE) / length(non_na_values) 
  ifelse(is.nan(prop) || is.na(prop), 0, prop)
})

df_feng_stats$prop_conserv_target <- apply(df_feng_glucOnly, 1, function(x) {
  non_na_values <- x[!is.na(x)]
  prop <- sum(x >= 70 & x <= 140, na.rm = TRUE) / length(non_na_values) 
  ifelse(is.nan(prop) || is.na(prop), 0, prop)
})

df_feng_stats$prop_above_target <- apply(df_feng_glucOnly, 1, function(x) {
  non_na_values <- x[!is.na(x)]
  prop <- sum(x >= 181 & x <= 250, na.rm = TRUE) / length(non_na_values) 
  ifelse(is.nan(prop) || is.na(prop), 0, prop)
})

df_feng_stats$prop_far_above_target <- apply(df_feng_glucOnly, 1, function(x) {
  non_na_values <- x[!is.na(x)]
  prop <- sum(x > 250, na.rm = TRUE) / length(non_na_values) 
  ifelse(is.nan(prop) || is.na(prop), 0, prop)
})
```

```{r}
require(dplyr)
# get list of columns to add, besides of the CGM statistics
cols_to_add = colnames(select(df_CGM_Char_HbA1c, -starts_with("Glucose")))

# get rid of non-feature columns and add the output A1c
df_feng_roc_ready = select(df_feng_roc_ready, -starts_with("Glucose"))
df_feng_stats_and_roc_ready = df_feng_roc_ready # to get only non-common columns
df_feng_roc_ready = cbind(df_feng_roc_ready, df_CGM_Char_HbA1c[,cols_to_add])
df_feng_roc_ready

df_feng_stats_ready = select(df_feng_stats, -starts_with("Glucose"))
df_feng_stats_ready = cbind(df_feng_stats_ready, df_CGM_Char_HbA1c[,cols_to_add])
df_feng_stats_ready

# create a df with both of the above merged
df_feng_stats_and_roc_ready = cbind(df_feng_stats_and_roc_ready, df_feng_stats_ready)
df_feng_stats_and_roc_ready
```


```{r}
set.seed(1)
# get random rows
train_feng = sample(1:nrow(df_feng_stats_and_roc_ready), 0.75*nrow(df_feng), replace = FALSE)
#train_feng

# get train and test set for all 3 datasets above
df_feng_roc_ready_train = df_feng_roc_ready[train_feng,]
df_feng_roc_ready_test = df_feng_roc_ready[-train_feng,]

df_feng_stats_ready_train = df_feng_stats_ready[train_feng,]
df_feng_stats_ready_test = df_feng_stats_ready[-train_feng,]

df_feng_stats_and_roc_ready_train = df_feng_stats_and_roc_ready[train_feng,]
df_feng_stats_and_roc_ready_test = df_feng_stats_and_roc_ready[-train_feng,]
```

```{r}
# Remove one test outlier with weird behaviors and only 104 readings
df_feng_roc_ready_test = df_feng_roc_ready_test[-173,]
df_feng_stats_ready_test = df_feng_stats_ready_test[-173,]
df_feng_stats_and_roc_ready_test = df_feng_stats_and_roc_ready_test[-173,]
```


```{r}
# Write handmade features as csv for further analyses
write.csv(df_feng_roc_ready_train, "df_feng_roc_ready_train.csv")
write.csv(df_feng_roc_ready_test, "df_feng_roc_ready_test.csv")

write.csv(df_feng_stats_ready_train, "df_feng_stats_ready_train.csv")
write.csv(df_feng_stats_ready_test, "df_feng_stats_ready_test.csv")

write.csv(df_feng_stats_and_roc_ready_train, "df_feng_stats_and_roc_ready_train.csv")
write.csv(df_feng_stats_and_roc_ready_test, "df_feng_stats_and_roc_ready_test.csv")
```


### Linear Regression

```{r}
# as part of replicating the study, we only keep the same statistics they had in the stats dataset:

df_feng_stats_ready_test$min_gluc = NULL
df_feng_stats_ready_test$max_gluc = NULL
df_feng_stats_ready_test$mad_gluc = NULL
df_feng_stats_ready_test$median_gluc = NULL
df_feng_stats_ready_test

df_feng_stats_ready_train$min_gluc = NULL
df_feng_stats_ready_train$max_gluc = NULL
df_feng_stats_ready_train$mad_gluc = NULL
df_feng_stats_ready_train$median_gluc = NULL
df_feng_stats_ready_train
```



Now some basic linear regressions to start.

```{r}
lm_feng_roc.1 = lm(data = df_feng_roc_ready_train, HbA1c ~ . - PtID - HbA1c_prev)
```
```{r}
lm_feng_stats.1 = lm(data = df_feng_stats_ready_train, HbA1c ~ . - PtID - HbA1c_prev)
```
```{r}
lm_feng_stats_and_roc.1 = lm(data = df_feng_stats_and_roc_ready_train, HbA1c ~ . - PtID - HbA1c_prev)
```

```{r}
summary(lm_feng_roc.1)
summary(lm_feng_stats.1)
summary(lm_feng_stats_and_roc.1)
```



```{r}
pred_feng_roc.1 = predict(lm_feng_roc.1, newdata = df_feng_roc_ready_test, 
                          na.action = na.pass)

pred_feng_stats.1 = predict(lm_feng_stats.1, newdata = df_feng_stats_ready_test, 
                            na.action = na.pass)

pred_feng_stats_and_roc.1 = predict(lm_feng_stats_and_roc.1, 
                                    newdata = df_feng_stats_and_roc_ready_test, 
                                    na.action = na.pass)
```


```{r}
plot(y = as.numeric(unlist(df_feng_roc_ready_test[,"HbA1c"])), x = pred_feng_roc.1)
abline(coef = c(0, 1))

plot(y = as.numeric(unlist(df_feng_stats_ready_test[,"HbA1c"])), x = pred_feng_stats.1)
abline(coef = c(0, 1))

plot(y = as.numeric(unlist(df_feng_stats_and_roc_ready_test[,"HbA1c"])), x = pred_feng_stats_and_roc.1)
abline(coef = c(0, 1))
```



```{r}
MSE_roc = mean((pred_feng_roc.1 - df_feng_roc_ready_test$HbA1c)^2)
MSE_roc
MSE_stats = mean((pred_feng_stats.1 - df_feng_stats_ready_test$HbA1c)^2)
MSE_stats
MSE_stats_and_roc = mean((pred_feng_stats_and_roc.1 - df_feng_stats_and_roc_ready_test$HbA1c)^2)
MSE_stats_and_roc
```

```{r}
RMSE_roc = sqrt(MSE_roc)
RMSE_roc
RMSE_stats = sqrt(MSE_stats)
RMSE_stats
RMSE_stats_and_roc = sqrt(MSE_stats_and_roc)
RMSE_stats_and_roc
```
```{r}
MAE_roc = mean(abs(pred_feng_roc.1 - df_feng_roc_ready_test$HbA1c))
MAE_roc
MAE_stats = mean(abs(pred_feng_stats.1 - df_feng_stats_ready_test$HbA1c))
MAE_stats
MAE_stats_and_roc = mean(abs(pred_feng_stats_and_roc.1 - df_feng_stats_and_roc_ready_test$HbA1c))
MAE_stats_and_roc
```

```{r}
debile_debile_feng.1 = rep(mean(df_feng_stats_and_roc_ready_test$HbA1c), nrow(df_feng_stats_and_roc_ready_test))
MSE_interecpt_only = mean((debile_debile_feng.1 - df_feng_stats_and_roc_ready_test$HbA1c)^2)
MAE_interecpt_only = mean(abs(debile_debile_feng.1 - df_feng_stats_and_roc_ready_test$HbA1c))
MSE_interecpt_only
MAE_interecpt_only
```

```{r}
which(rowSums(is.na(data.matrix(select(df_feng_roc_ready_train, -c("HbA1c", "HbA1c_prev", "PtID"))))) > 0)
```
```{r}
which(is.na(df_feng_roc_ready_train))
```
```{r}
df_feng_roc_ready_train[df_feng_roc_ready_train$PtID==432, ]
```


Now fitting some Ridge and Lasso :

```{r}
require(glmnet)

### Ridge ###

lm_feng_roc.ridge = cv.glmnet(data.matrix(select(df_feng_roc_ready_train, 
                                                 -c("HbA1c", "HbA1c_prev", "PtID"))), 
                              df_feng_roc_ready_train$HbA1c, alpha = 0)

lm_feng_stats.ridge = cv.glmnet(data.matrix(select(df_feng_stats_ready_train, 
                                                   -c("HbA1c", "HbA1c_prev", "PtID"))), 
                              df_feng_stats_ready_train$HbA1c, alpha = 0)

lm_feng_stats_and_roc.ridge = cv.glmnet(data.matrix(select(
  df_feng_stats_and_roc_ready_train, -
    c("HbA1c", "HbA1c_prev", "PtID"))), 
  df_feng_stats_and_roc_ready_train$HbA1c, alpha = 0)

### Lasso ###

lm_feng_roc.lasso = cv.glmnet(data.matrix(select(df_feng_roc_ready_train, 
                                                 -c("HbA1c", "HbA1c_prev", "PtID"))), 
                              df_feng_roc_ready_train$HbA1c, alpha = 1)

lm_feng_stats.lasso = cv.glmnet(data.matrix(select(df_feng_stats_ready_train, 
                                                   -c("HbA1c", "HbA1c_prev", "PtID"))), 
                              df_feng_stats_ready_train$HbA1c, alpha = 1)

lm_feng_stats_and_roc.lasso = cv.glmnet(data.matrix(select(
  df_feng_stats_and_roc_ready_train, 
  -c("HbA1c", "HbA1c_prev", "PtID"))), 
  df_feng_stats_and_roc_ready_train$HbA1c, alpha = 1)

```



```{r}
### Ridge ###

plot(lm_feng_roc.ridge)

plot(lm_feng_stats.ridge)

plot(lm_feng_stats_and_roc.ridge)


### Lasso ###

plot(lm_feng_roc.lasso)

plot(lm_feng_stats.lasso)

plot(lm_feng_stats_and_roc.lasso)
```




```{r}
### Ridge ###

bestlambda_roc.ridge = lm_feng_roc.ridge$lambda.min # value of lambda which gives the smaller cross-validation MSE

bestlambda_stats.ridge = lm_feng_stats.ridge$lambda.min

bestlambda_stats_and_roc.ridge = lm_feng_stats_and_roc.ridge$lambda.min

### Lasso ###

bestlambda_roc.lasso = lm_feng_roc.lasso$lambda.min # value of lambda which gives the smaller cross-validation MSE

oneSElambda_roc.lasso = lm_feng_roc.lasso$lambda.1se

bestlambda_stats.lasso = lm_feng_stats.lasso$lambda.min

oneSElambda_stats.lasso = lm_feng_stats.lasso$lambda.1se

bestlambda_stats_and_roc.lasso = lm_feng_stats_and_roc.lasso$lambda.min

oneSElambda_stats_and_roc.lasso = lm_feng_stats_and_roc.lasso$lambda.1se
```


```{r}
### Ridge ### 

pred_feng_roc.ridge = predict(lm_feng_roc.ridge, s=bestlambda_roc.ridge, 
                              newx = data.matrix(select(df_feng_roc_ready_test, 
                                                        -c("HbA1c", "HbA1c_prev", "PtID"))))

pred_feng_stats.ridge = predict(lm_feng_stats.ridge, s=bestlambda_stats.ridge, 
                              newx = data.matrix(select(df_feng_stats_ready_test, 
                                                        -c("HbA1c", "HbA1c_prev", "PtID"))))

pred_feng_stats_and_roc.ridge = predict(lm_feng_stats_and_roc.ridge, 
                                        s=bestlambda_stats_and_roc.ridge, 
                                        newx = data.matrix(select(
                                          df_feng_stats_and_roc_ready_test, 
                                          -c("HbA1c", "HbA1c_prev", "PtID"))))

### Lasso ###

# Using best Lambda

pred_feng_roc.lasso = predict(lm_feng_roc.lasso, s=bestlambda_roc.lasso, 
                              newx = data.matrix(select(df_feng_roc_ready_test, 
                                                        -c("HbA1c", "HbA1c_prev", "PtID"))))

pred_feng_stats.lasso = predict(lm_feng_stats.lasso, s=bestlambda_stats.lasso, 
                              newx = data.matrix(select(df_feng_stats_ready_test, 
                                                        -c("HbA1c", "HbA1c_prev", "PtID"))))

pred_feng_stats_and_roc.lasso = predict(lm_feng_stats_and_roc.lasso, 
                                        s=bestlambda_stats_and_roc.lasso, 
                                        newx = data.matrix(select(
                                          df_feng_stats_and_roc_ready_test, 
                                          -c("HbA1c", "HbA1c_prev", "PtID"))))

# Using One-SE Lambda

pred_feng_roc.lasso.1se = predict(lm_feng_roc.lasso, s=oneSElambda_roc.lasso, 
                              newx = data.matrix(select(df_feng_roc_ready_test, 
                                                        -c("HbA1c", "HbA1c_prev", "PtID"))))

pred_feng_stats.lasso.1se = predict(lm_feng_stats.lasso, s=oneSElambda_stats.lasso, 
                              newx = data.matrix(select(df_feng_stats_ready_test, 
                                                        -c("HbA1c", "HbA1c_prev", "PtID"))))

pred_feng_stats_and_roc.lasso.1se = predict(lm_feng_stats_and_roc.lasso, 
                                        s=oneSElambda_stats_and_roc.lasso, 
                                        newx = data.matrix(select(
                                          df_feng_stats_and_roc_ready_test, 
                                          -c("HbA1c", "HbA1c_prev", "PtID"))))

# Lasso coefficients kept by best lambda
coef_feng_roc.lasso = predict(lm_feng_roc.lasso, type="coefficients", s=bestlambda_roc.lasso)
coef_feng_roc.lasso

coef_feng_stats.lasso = predict(lm_feng_stats.lasso, type="coefficients", s=bestlambda_stats.lasso)
coef_feng_stats.lasso

coef_feng_stats_and_roc.lasso = predict(lm_feng_stats_and_roc.lasso, type="coefficients", s=bestlambda_stats_and_roc.lasso)
coef_feng_stats_and_roc.lasso

# Lasso coefficients kept by One-SE lambda
coef_feng_roc.lasso.1se = predict(lm_feng_roc.lasso, type="coefficients", s=oneSElambda_roc.lasso)
coef_feng_roc.lasso.1se

coef_feng_stats.lasso.1se = predict(lm_feng_stats.lasso, type="coefficients", s=oneSElambda_stats.lasso)
coef_feng_stats.lasso.1se

coef_feng_stats_and_roc.lasso.1se = predict(lm_feng_stats_and_roc.lasso, type="coefficients", s=oneSElambda_stats_and_roc.lasso)
coef_feng_stats_and_roc.lasso.1se
```



```{r}
# Get the variable names of the non-zero coefficients for lasso
var_feng_roc.lasso = row.names(coef_feng_roc.lasso)[which(coef_feng_roc.lasso != 0)]
var_feng_stats.lasso = row.names(coef_feng_stats.lasso)[which(coef_feng_stats.lasso != 0)]
var_feng_stats_and_roc.lasso = row.names(coef_feng_stats_and_roc.lasso)[which(coef_feng_stats_and_roc.lasso != 0)]

cat("Variables kept by Lasso in roc:", var_feng_roc.lasso, "\n")
cat("Variables kept by Lasso in stats:", var_feng_stats.lasso, "\n")
cat("Variables kept by Lasso in stats_and_roc:", var_feng_stats_and_roc.lasso, "\n")

var_feng_roc.lasso.1se = row.names(coef_feng_roc.lasso.1se)[which(coef_feng_roc.lasso.1se != 0)]
var_feng_stats.lasso.1se = row.names(coef_feng_stats.lasso.1se)[which(coef_feng_stats.lasso.1se != 0)]
var_feng_stats_and_roc.lasso.1se = row.names(coef_feng_stats_and_roc.lasso.1se)[which(coef_feng_stats_and_roc.lasso.1se != 0)]

cat("Variables kept by One-SE Lasso in roc:", var_feng_roc.lasso.1se, "\n")
cat("Variables kept by One-SE Lasso in stats:", var_feng_stats.lasso.1se, "\n")
cat("Variables kept by One-SE Lasso in stats_and_roc:", var_feng_stats_and_roc.lasso.1se, "\n")
```


```{r}
### Ridge ###

plot(y = as.numeric(unlist(df_feng_roc_ready_test[,"HbA1c"])), x = pred_feng_roc.ridge)
abline(coef = c(0, 1))

plot(y = as.numeric(unlist(df_feng_stats_ready_test[,"HbA1c"])), x = pred_feng_stats.ridge)
abline(coef = c(0, 1))

plot(y = as.numeric(unlist(df_feng_stats_and_roc_ready_test[,"HbA1c"])), x = pred_feng_stats_and_roc.ridge)
abline(coef = c(0, 1))

### Lasso ### 

# Best Lambda

plot(y = as.numeric(unlist(df_feng_roc_ready_test[,"HbA1c"])), x = pred_feng_roc.lasso)
abline(coef = c(0, 1))

plot(y = as.numeric(unlist(df_feng_stats_ready_test[,"HbA1c"])), x = pred_feng_stats.lasso)
abline(coef = c(0, 1))

plot(y = as.numeric(unlist(df_feng_stats_and_roc_ready_test[,"HbA1c"])), x = pred_feng_stats_and_roc.lasso)
abline(coef = c(0, 1))

# One-SE Lambda

plot(y = as.numeric(unlist(df_feng_roc_ready_test[,"HbA1c"])), x = pred_feng_roc.lasso.1se)
abline(coef = c(0, 1))

plot(y = as.numeric(unlist(df_feng_stats_ready_test[,"HbA1c"])), x = pred_feng_stats.lasso.1se)
abline(coef = c(0, 1))

plot(y = as.numeric(unlist(df_feng_stats_and_roc_ready_test[,"HbA1c"])), 
     x = pred_feng_stats_and_roc.lasso.1se)
abline(coef = c(0, 1))
```




```{r}
### Ridge ### 

MSE_roc.ridge = mean((pred_feng_roc.ridge - df_feng_roc_ready_test$HbA1c)^2)
MSE_roc.ridge
MSE_stats.ridge = mean((pred_feng_stats.ridge - df_feng_stats_ready_test$HbA1c)^2)
MSE_stats.ridge
MSE_stats_and_roc.ridge = mean((pred_feng_stats_and_roc.ridge - df_feng_stats_and_roc_ready_test$HbA1c)^2)
MSE_stats_and_roc.ridge

### Lasso ###

# Best Lambda 

MSE_roc.lasso = mean((pred_feng_roc.lasso - df_feng_roc_ready_test$HbA1c)^2)
MSE_roc.lasso
MSE_stats.lasso = mean((pred_feng_stats.lasso - df_feng_stats_ready_test$HbA1c)^2)
MSE_stats.lasso
MSE_stats_and_roc.lasso = mean((pred_feng_stats_and_roc.lasso - df_feng_stats_and_roc_ready_test$HbA1c)^2)
MSE_stats_and_roc.lasso

# One-SE Lambda 

MSE_roc.lasso.1se = mean((pred_feng_roc.lasso.1se - df_feng_roc_ready_test$HbA1c)^2)
MSE_roc.lasso.1se
MSE_stats.lasso.1se = mean((pred_feng_stats.lasso.1se - df_feng_stats_ready_test$HbA1c)^2)
MSE_stats.lasso.1se
MSE_stats_and_roc.lasso.1se = mean((pred_feng_stats_and_roc.lasso.1se - df_feng_stats_and_roc_ready_test$HbA1c)^2)
MSE_stats_and_roc.lasso.1se
```

```{r}
### Ridge ###

MAE_roc.ridge = mean(abs(pred_feng_roc.ridge - df_feng_roc_ready_test$HbA1c))
MAE_roc.ridge
MAE_stats.ridge = mean(abs(pred_feng_stats.ridge - df_feng_stats_ready_test$HbA1c))
MAE_stats.ridge
MAE_stats_and_roc.ridge = mean(abs(pred_feng_stats_and_roc.ridge - df_feng_stats_and_roc_ready_test$HbA1c))
MAE_stats_and_roc.ridge

### Lasso ###

# Best Lambda

MAE_roc.lasso = mean(abs(pred_feng_roc.lasso - df_feng_roc_ready_test$HbA1c))
MAE_roc.lasso
MAE_stats.lasso = mean(abs(pred_feng_stats.lasso - df_feng_stats_ready_test$HbA1c))
MAE_stats.lasso
MAE_stats_and_roc.lasso = mean(abs(pred_feng_stats_and_roc.lasso - df_feng_stats_and_roc_ready_test$HbA1c))
MAE_stats_and_roc.lasso

# One-SE Lambda

MAE_roc.lasso.1se = mean(abs(pred_feng_roc.lasso.1se - df_feng_roc_ready_test$HbA1c))
MAE_roc.lasso
MAE_stats.lasso.1se = mean(abs(pred_feng_stats.lasso.1se - df_feng_stats_ready_test$HbA1c))
MAE_stats.lasso.1se
MAE_stats_and_roc.lasso.1se = mean(abs(pred_feng_stats_and_roc.lasso.1se - df_feng_stats_and_roc_ready_test$HbA1c))
MAE_stats_and_roc.lasso.1se
```


Now let's build the null model to compare performance

```{r}
lm_feng.null = lm(data = df_feng_stats_and_roc_ready_test, HbA1c ~ 1)
summary(lm_feng.null)
pred_feng.null = predict(lm_feng.null, newdata = df_feng_stats_and_roc_ready_test)
MSE.null = mean((pred_feng.null - df_feng_stats_and_roc_ready_test$HbA1c)^2)
MAE.null = mean(abs(pred_feng.null - df_feng_stats_and_roc_ready_test$HbA1c))
```


```{r}
plot_prediction <- function(pred_val) {
  df_temp = data.frame(index = as.integer(row.names(df_feng_roc_ready_test)), 
                           pred_val = as.numeric(pred_val), 
                           true_val = as.numeric(unlist(df_feng_roc_ready_test[,"HbA1c"])))
  df_temp = df_temp[order(df_temp$true_val), ]
  ggplot(data = df_temp, aes(x = index)) +
    geom_line(aes(y = pred_val), color = "firebrick", linewidth = 1.5, alpha = 0.7) +
    geom_line(aes(y = true_val), color = "lightblue", linewidth = 1.5, alpha = 0.7)
}
plot_prediction_2 <- function(pred_val) {
  df_temp = data.frame(pred_val = as.numeric(pred_val), 
                       true_val = as.numeric(unlist(df_feng_roc_ready_test[,"HbA1c"])))
  df_temp = df_temp[order(df_temp$true_val), ]
  ggplot(data = df_temp, aes(x = 1:length(df_temp$true_val))) +
    geom_line(aes(y = pred_val), color = "firebrick", linewidth = 1.5, alpha = 0.7) +
    geom_line(aes(y = true_val), color = "lightblue", linewidth = 1.5, alpha = 0.7)
}

```

```{r}
plot_prediction_2(pred_feng_roc.1)
plot_prediction_2(pred_feng_roc.ridge)
plot_prediction_2(pred_feng_roc.lasso)
plot_prediction_2(pred_feng_roc.lasso.1se)
plot_prediction_2(pred_feng_stats.1)
plot_prediction_2(pred_feng_stats.ridge)
plot_prediction_2(pred_feng_stats.lasso)
plot_prediction_2(pred_feng_stats.lasso.1se)
plot_prediction_2(pred_feng_stats_and_roc.1)
plot_prediction_2(pred_feng_stats_and_roc.ridge)
plot_prediction_2(pred_feng_stats_and_roc.lasso)
plot_prediction_2(pred_feng_stats_and_roc.lasso.1se)
plot_prediction_2(pred_feng.null)
```

```{r}
df_CGM_Char_HbA1c[df_CGM_Char_HbA1c$PtID == 386, ]
```


```{r}
df_feng_stats_and_roc_ready[df_feng_stats_and_roc_ready$PtID == 386,]
```


```{r}
plot_prediction(pred_feng_roc.1)
plot_prediction(pred_feng_roc.ridge)
plot_prediction(pred_feng_roc.lasso)
plot_prediction(pred_feng_roc.lasso.1se)
plot_prediction(pred_feng_stats.1)
plot_prediction(pred_feng_stats.ridge)
plot_prediction(pred_feng_stats.lasso)
plot_prediction(pred_feng_stats.lasso.1se)
plot_prediction(pred_feng_stats_and_roc.1)
plot_prediction(pred_feng_stats_and_roc.ridge)
plot_prediction(pred_feng_stats_and_roc.lasso)
plot_prediction(pred_feng_stats_and_roc.lasso.1se)
plot_prediction(pred_feng.null)
```




```{r}
# Null


within_1_error.null = (1 - sum(abs(pred_feng.null - df_feng_stats_and_roc_ready_test$HbA1c) >= 1)/nrow(df_feng_stats_and_roc_ready_test))*100


within_0.5_error.null = (1 - sum(abs(pred_feng.null - df_feng_stats_and_roc_ready_test$HbA1c) >= 0.5)/nrow(df_feng_stats_and_roc_ready_test))*100


# OLS

within_1_error_roc.1 = (1 - sum(abs(pred_feng_roc.1 - df_feng_roc_ready_test$HbA1c) >= 1)/nrow(df_feng_stats_and_roc_ready_test))*100

within_1_error_stats.1 = (1 - sum(abs(pred_feng_stats.1 - df_feng_stats_ready_test$HbA1c) >= 1)/nrow(df_feng_stats_and_roc_ready_test))*100

within_1_error_stats_and_roc.1 = (1 - sum(abs(pred_feng_stats_and_roc.1 - df_feng_stats_and_roc_ready_test$HbA1c) >= 1)/nrow(df_feng_stats_and_roc_ready_test))*100


within_0.5_error_roc.1 = (1 - sum(abs(pred_feng_roc.1 - df_feng_roc_ready_test$HbA1c) >= 0.5)/nrow(df_feng_stats_and_roc_ready_test))*100

within_0.5_error_stats.1 = (1 - sum(abs(pred_feng_stats.1 - df_feng_stats_ready_test$HbA1c) >= 0.5)/nrow(df_feng_stats_and_roc_ready_test))*100

within_0.5_error_stats_and_roc.1 = (1 - sum(abs(pred_feng_stats_and_roc.1 - df_feng_stats_and_roc_ready_test$HbA1c) >= 0.5)/nrow(df_feng_stats_and_roc_ready_test))*100

# Ridge

within_1_error_roc.ridge = (1 - sum(abs(pred_feng_roc.ridge - df_feng_roc_ready_test$HbA1c) >= 1)/nrow(df_feng_stats_and_roc_ready_test))*100

within_1_error_stats.ridge = (1 - sum(abs(pred_feng_stats.ridge - df_feng_stats_ready_test$HbA1c) >= 1)/nrow(df_feng_stats_and_roc_ready_test))*100

within_1_error_stats_and_roc.ridge = (1 - sum(abs(pred_feng_stats_and_roc.ridge - df_feng_stats_and_roc_ready_test$HbA1c) >= 1)/nrow(df_feng_stats_and_roc_ready_test))*100


within_0.5_error_roc.ridge = (1 - sum(abs(pred_feng_roc.ridge - df_feng_roc_ready_test$HbA1c) >= 0.5)/nrow(df_feng_stats_and_roc_ready_test))*100

within_0.5_error_stats.ridge = (1 - sum(abs(pred_feng_stats.ridge - df_feng_stats_ready_test$HbA1c) >= 0.5)/nrow(df_feng_stats_and_roc_ready_test))*100

within_0.5_error_stats_and_roc.ridge = (1 - sum(abs(pred_feng_stats_and_roc.ridge - df_feng_stats_and_roc_ready_test$HbA1c) >= 0.5)/nrow(df_feng_stats_and_roc_ready_test))*100

# Lasso

# Best Lambda

within_1_error_roc.lasso = (1 - sum(abs(pred_feng_roc.lasso - df_feng_roc_ready_test$HbA1c) >= 1)/nrow(df_feng_stats_and_roc_ready_test))*100

within_1_error_stats.lasso = (1 - sum(abs(pred_feng_stats.lasso - df_feng_stats_ready_test$HbA1c) >= 1)/nrow(df_feng_stats_and_roc_ready_test))*100

within_1_error_stats_and_roc.lasso = (1 - sum(abs(pred_feng_stats_and_roc.lasso - df_feng_stats_and_roc_ready_test$HbA1c) >= 1)/nrow(df_feng_stats_and_roc_ready_test))*100


within_0.5_error_roc.lasso = (1 - sum(abs(pred_feng_roc.lasso - df_feng_roc_ready_test$HbA1c) >= 0.5)/nrow(df_feng_stats_and_roc_ready_test))*100

within_0.5_error_stats.lasso = (1 - sum(abs(pred_feng_stats.lasso - df_feng_stats_ready_test$HbA1c) >= 0.5)/nrow(df_feng_stats_and_roc_ready_test))*100

within_0.5_error_stats_and_roc.lasso = (1 - sum(abs(pred_feng_stats_and_roc.lasso - df_feng_stats_and_roc_ready_test$HbA1c) >= 0.5)/nrow(df_feng_stats_and_roc_ready_test))*100

# One-SE Lambda

within_1_error_roc.lasso.1se = (1 - sum(abs(pred_feng_roc.lasso.1se - df_feng_roc_ready_test$HbA1c) >= 1)/nrow(df_feng_stats_and_roc_ready_test))*100

within_1_error_stats.lasso.1se = (1 - sum(abs(pred_feng_stats.lasso.1se - df_feng_stats_ready_test$HbA1c) >= 1)/nrow(df_feng_stats_and_roc_ready_test))*100

within_1_error_stats_and_roc.lasso.1se = (1 - sum(abs(pred_feng_stats_and_roc.lasso.1se - df_feng_stats_and_roc_ready_test$HbA1c) >= 1)/nrow(df_feng_stats_and_roc_ready_test))*100


within_0.5_error_roc.lasso.1se = (1 - sum(abs(pred_feng_roc.lasso.1se - df_feng_roc_ready_test$HbA1c) >= 0.5)/nrow(df_feng_stats_and_roc_ready_test))*100

within_0.5_error_stats.lasso.1se = (1 - sum(abs(pred_feng_stats.lasso.1se - df_feng_stats_ready_test$HbA1c) >= 0.5)/nrow(df_feng_stats_and_roc_ready_test))*100

within_0.5_error_stats_and_roc.lasso.1se = (1 - sum(abs(pred_feng_stats_and_roc.lasso.1se - df_feng_stats_and_roc_ready_test$HbA1c) >= 0.5)/nrow(df_feng_stats_and_roc_ready_test))*100
```


```{r}
require(ggplot2)

# Create a data frame with the category, subcategory, and within_1_error values
data = data.frame(
  model = rep(c("OLS", "Ridge", "Lasso - Best Lambda", "Lasso - One-SE Lambda"), each = 3),
  dataset = rep(c("roc", "stats", "stats_and_roc"), times = 4),
  within_1_error = c(
    within_1_error_roc.1, within_1_error_stats.1, within_1_error_stats_and_roc.1,
    within_1_error_roc.ridge, within_1_error_stats.ridge, within_1_error_stats_and_roc.ridge,
    within_1_error_roc.lasso, within_1_error_stats.lasso, within_1_error_stats_and_roc.lasso, 
    within_1_error_roc.lasso.1se, within_1_error_stats.lasso.1se, 
    within_1_error_stats_and_roc.lasso.1se),
  within_0.5_error = c(
    within_0.5_error_roc.1, within_0.5_error_stats.1, within_0.5_error_stats_and_roc.1,
    within_0.5_error_roc.ridge, within_0.5_error_stats.ridge, within_0.5_error_stats_and_roc.ridge,
    within_0.5_error_roc.lasso, within_0.5_error_stats.lasso, within_0.5_error_stats_and_roc.lasso,
    within_0.5_error_roc.lasso.1se, within_0.5_error_stats.lasso.1se, 
    within_0.5_error_stats_and_roc.lasso.1se)
)

# Plot bar chart using ggplot
ggplot(data, aes(x = factor(model, levels = c("OLS", "Ridge", "Lasso - Best Lambda", "Lasso - One-SE Lambda")), y = within_1_error, fill = dataset)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) +
  labs(x = "Model", y = "Within 1 Error (%)", fill = "Data Set Used") +
  ylim(0, 100) +
  theme_minimal()

ggplot(data, aes(x = factor(model, levels = c("OLS", "Ridge", "Lasso - Best Lambda", "Lasso - One-SE Lambda")), y = within_0.5_error, fill = dataset)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) +
  labs(x = "Model", y = "Within 0.5 Error (%)", fill = "Data Set Used") + 
  ylim(0, 100) +
  theme_minimal()


```



```{r}
require(knitr)

metrics <- data.frame(
  Model = c(
    "Null", "ROC - Classic", "ROC - Ridge", "ROC - Lasso - Best Lambda", 
    "ROC - Lasso - One-SE Lambda", 
    "Stats - Classic", "Stats - Ridge", "Stats - Lasso - Best Lambda", 
    "Stats - Lasso - One-SE Lambda", 
    "Full - Classic", "Full - Ridge", "Full - Lasso - Best Lambda", 
    "Full - Lasso - One-SE Lambda"
  ),
  MSE = round(c(
    MSE.null, MSE_roc, MSE_roc.ridge, MSE_roc.lasso, MSE_roc.lasso.1se, MSE_stats, MSE_stats.ridge,
    MSE_stats.lasso, MSE_stats.lasso.1se, MSE_stats_and_roc, MSE_stats_and_roc.ridge, 
    MSE_stats_and_roc.lasso, MSE_stats_and_roc.lasso.1se
  ), 3),
  MAE = round(c(
    MAE.null, MAE_roc, MAE_roc.ridge, MAE_roc.lasso, MAE_roc.lasso.1se, MAE_stats, MAE_stats.ridge,
    MAE_stats.lasso, MAE_stats.lasso.1se, MAE_stats_and_roc, MAE_stats_and_roc.ridge, 
    MAE_stats_and_roc.lasso, MAE_stats_and_roc.lasso.1se
  ), 3),
  Within_1_Error = round(c(
    within_1_error.null, within_1_error_roc.1, within_1_error_roc.ridge, within_1_error_roc.lasso, 
    within_1_error_roc.lasso.1se, within_1_error_stats.1, within_1_error_stats.ridge, 
    within_1_error_stats.lasso, within_1_error_stats.lasso.1se, within_1_error_stats_and_roc.1, 
    within_1_error_stats_and_roc.ridge, within_1_error_stats_and_roc.lasso, 
    within_1_error_stats_and_roc.lasso.1se
  ), 2),
  Within_0.5_Error = round(c(
    within_0.5_error.null, within_0.5_error_roc.1, within_0.5_error_roc.ridge, 
    within_0.5_error_roc.lasso, within_0.5_error_roc.lasso.1se, within_0.5_error_stats.1, 
    within_0.5_error_stats.ridge, within_0.5_error_stats.lasso, within_0.5_error_stats.lasso.1se,
    within_0.5_error_stats_and_roc.1, within_0.5_error_stats_and_roc.ridge, 
    within_0.5_error_stats_and_roc.lasso, within_0.5_error_stats_and_roc.lasso.1se
  ), 2)
)

kable(metrics)
````


### Now some additionals regressions, more specific

```{r}
round(cor(select_if(df_feng_stats_and_roc_ready_train, is.numeric)), 2)[,"HbA1c"]
```
We see that the most correlated variables are: mean_gluc, sd_gluc, median_gluc, mad_gluc, prop_target (TIR), pro_converv_target, prop_above_target, prop_far_above_target (and obviously HbA1c_prev).


```{r}
simple_lm <- function(train, test, output, input) {
  # run lm model and summarize
  formula = as.formula(paste(output, "~", paste(input, collapse = "+")))
  temp_lm = lm(data = train, formula = formula)
  print(summary(temp_lm))
  cat("\n \n ################################################## \n \n \n")
  
  # predict 
  pred_temp = predict(temp_lm, newdata = test, na.action = na.pass)
  
  # compute performance metrics
  within_1_percent_value = (1 - sum(abs(pred_temp - test[[output]]) >= 1)/nrow(test))*100
  within_0.5_percent_value = (1 - sum(abs(pred_temp - test[[output]]) >= 0.5)/nrow(test))*100
  mae_value = mean(abs(pred_temp - test[[output]]))
  mse_value = mean((pred_temp - test[[output]])^2)
  cat("MSE:", mse_value, "\n")
  cat("MAE:", mae_value, "\n")
  cat("Within 1% error:", within_1_percent_value, "\n")
  cat("Within 0.5% error:", within_0.5_percent_value, "\n")

  
  cat("\n \n ################################################## \n \n \n")

  plot(y = test[[output]], x = pred_temp, col = "black")
  abline(coef = c(0, 1), col = "firebrick", lwd = 1.5)
  abline(coef = c(1, 1), col = "lightgray")
  abline(coef = c(-1, 1), col = "lightgray")
  abline(coef = c(0.5, 1), col = "darkgray")
  abline(coef = c(-0.5, 1), col = "darkgray")

  par(mfrow = c(2, 1))
  plot_prediction_2(pred_temp)
  
  
}
```

```{r}
simple_lm(df_feng_stats_and_roc_ready_train, df_feng_stats_and_roc_ready_test, "HbA1c", "prop_target")
```
```{r}
simple_lm(df_feng_stats_and_roc_ready_train, df_feng_stats_and_roc_ready_test, "HbA1c", "prop_conserv_target")
```

```{r}
simple_lm(df_feng_stats_and_roc_ready_train, df_feng_stats_and_roc_ready_test, "HbA1c", "mean_gluc")
```

```{r}
simple_lm(df_feng_stats_and_roc_ready_train, df_feng_stats_and_roc_ready_test, "HbA1c", "prop_clini_hypo")
```

```{r}
simple_lm(df_feng_stats_and_roc_ready_train, df_feng_stats_and_roc_ready_test, "HbA1c", c("AgeAsOfRandDt"))
```

```{r warning = F}
simple_lm(df_feng_stats_and_roc_ready_train, df_feng_stats_and_roc_ready_test, "HbA1c", "AgeAsOfRandDt * mean_gluc * Race")
```

```{r}
simple_lm(df_feng_stats_and_roc_ready_train, df_feng_stats_and_roc_ready_test, "HbA1c", "AgeAsOfRandDt * mean_gluc")
```


```{r}
simple_lm(df_feng_stats_and_roc_ready_train, df_feng_stats_and_roc_ready_test, "HbA1c", c("prop_target", "mean_gluc"))
```


```{r}
simple_lm(df_feng_stats_and_roc_ready_train, df_feng_stats_and_roc_ready_test, "HbA1c", 
          "mean_gluc * prop_target * sd_gluc * prop_above_target * prop_far_above_target")
```






### some plots

```{r}
hist(df_feng_stats_and_roc_ready_train[["HbA1c"]], breaks = 75)
hist(df_feng_stats_and_roc_ready_test[["HbA1c"]], breaks = 25)
```

```{r}
plot(y = df_feng_stats_and_roc_ready_train$HbA1c, x = df_feng_stats_and_roc_ready_train$mean_gluc)
```

```{r}
plot(select_if(df_feng_stats_and_roc_ready_train, is.numeric)[,c(14:19, 25)])
```

```{r}
plot(select_if(df_feng_stats_and_roc_ready_train, is.numeric)[,c(7, 8, 14:19, 25)])
```

### Interesting associations:

mean and sd gluc:

```{r}
plot(df_feng_stats_and_roc_ready_train$mean_gluc, df_feng_stats_and_roc_ready_train$sd_gluc)
```


